<!-- README.md is generated from README.Rmd. Please edit that file -->

# rpkgkit

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/rpkgkit)](https://CRAN.R-project.org/package=rpkgkit)
[![R-CMD-check](https://github.com/Exceret/rpkgkit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Exceret/rpkgkit/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/Exceret/rpkgkit/graph/badge.svg)](https://app.codecov.io/gh/Exceret/rpkgkit)
<!-- badges: end -->

The goal of rpkgkit is to provide useful functions for R package development.

## Features

All functions can detect active file contexts in RStudio and Positron, so generally file path can be omitted.

### Code Formatting & Linting

Require extra installation:

- `air_format()` - Format R code using [air](https://github.com/posit-dev/air)
  
- `flir_fix()` - Automatically fix linting issues using [flir](https://github.com/etiennebacher/flir)

### Available Standalone Scripts

Use `usethis::use_standalone("Exceret/rpkgkit", "<name>")` to import:

- args_to_func.R: matching arguments to function calls:

```{r}
f1 <- function(a, b) a + b
f2 <- function(x, y, ...) x * y
f3 <- function(p, q) p - q

args <- list(a = 1, b = 2)

# Strict matching (default): returns f1 only
match_func_to_args(args, f1, f2, f3)
```

- caller_cli.R: show where the cli function is called from

```{r}
decorated <- add_caller_to_cli(cli::cli_alert_info)
foo1 <- \() {
  print("I' m in foo1")
  decorated("<- where is this called from?")
}
foo2 <- \() {
  print("I' m in foo2")
  foo1()
}
bar <- function() {
  print("I' m in bar")
  foo2()
}
bar()
# [1] "I' m in bar"
# [1] "I' m in foo2"
# [1] "I' m in foo1"
# ℹ [foo1()]: <- where is this called from?
```

- colorful_cli.R: Easier color calling in one cli function

```{r}
color_cli <- create_colorful_cli_env()
color_cli$cli_alert_danger("{.red This is a red message}")
color_cli$cli_alert_info("{.blue This is a blue message}")
color_cli$cli_alert_info("{.orange This is an orange message}")
# cyan, green, magenta, yellow, purple, etc

color_cli2 <- create_colorful_cli_env(cli_theme = generate_color_theme()) # more color but slower
color_cli2$cli_alert_success(
  "{.violetred3 R}{.orange a}{.yellow i}{.green n}{.cyan b}{.blue o}{.purple w}"
)
```

- match_arg.R: partial matching of arguments to function calls, like `match.arg`, `rlang::arg_match`
  
- ts_cli.R: timestamp cli function

```{r}
ts_cli <- create_ts_cli_env()
ts_cli$cli_alert_info("Hello, world!")
# ℹ [2026/05/30 22:45:42] Hello, world!
```

### Standalone File Management

- `inquire_standalone()` - List standalone files available in a GitHub repository's `R/` directory

- `create_standalone()` - Create standalone utility files in your package

- `update_time_in_standalone()` - Update `last-updated` field in standalone files

- `add_changelog_in_standalone()` - Add changelog entries to standalone files

### NEWS.md Management

- `news_md_add_entry()` - Add new entries to NEWS.md following CRAN guidelines
  
- `news_md_check()` - Validate NEWS.md format for CRAN compliance

- `news_md_show()` - Display NEWS.md content with optional filtering

### Other Utilities

- `make_func_call_explicit()` - Make function calls explicit by adding package prefixes

- `render_rmd()` - Render R Markdown documents
  
- `use_hexsticker()` - Paste hex sticker to README.md

